Co-moving stars in \textsl{Gaia DR1}

### 2016-10-03 paper/ms.tex
words in data section

% define macros for text
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\figname}{Figure}
\newcommand{\eqname}{Equation}

% define macros for math
\newcommand{\given}{\,|\,}
\newcommand{\normal}{{\mathcal{N}}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transp}[1]{{#1}^{\!\mathsf{T}}}
\newcommand{\inv}[1]{{#1}^{-1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\vperp}{\bs{v}^\perp}
\newcommand{\propm}{\bs{\mu}}
\newcommand{\matrx}[1]{\mathbf{#1}}
\newcommand{\kms}{\rm km~s^{-1}}
\newcommand{\data}{\mathrm{data}}

@@

\title{Co-moving stars in \textsl{Gaia DR1}}
\author{Adrian M. Price-Whelan\altaffilmark{\pu,\adrn},
        Semyeong Oh\altaffilmark{\pu},
        Timothy D. Morton\altaffilmark{\pu},
        David W. Hogg\altaffilmark{\ccpp,\mpia},
        David N. Spergel\altaffilmark{\pu,\cca}
}

@@

@@

@@

\appendix

Here we derive analytic expressions for the marginal likelihood integrals above.

\todo{Hypothesis 1}

The integrals for Hypothesis 2 are somewhat easier because the true velocity
vector can be considered to be already in the orthogonal basis defined by the
tangent plane at the coordinates of the source (we are assuming an isotropic
velocity prior). We will analytically integrate over the velocity components
$(v_\alpha,v_\delta,v_r)$ and leave the distance integral to be done
numerically:
\begin{equation}
  \int \, \dd^3\bs{v} \, L(\bs{v}, d) \, p(\bs{v})
\end{equation}
The covariance matrix over parallax, proper motion components, and
radial velocity (if they exist) is $\matrx{C}$ and the inverse variance matrix
is $\matrx{V}$. Defining the constant
\begin{equation}
  \mathcal{Z} = \left[ \det \left(\frac{\matrx{V}}{2\pi} \right) \right]^{1/2}
\end{equation}
the integrand can be written
\begin{equation}
  L(\bs{v}, d) \, p(\bs{v}) = \mathcal{Z} \,
    \exp \left[ -\frac{1}{2} \transp{\left(\bs{x} - \bs{x}_\theta \right)} \,
    \matrx{V} \,
    \left(\bs{x} - \bs{x}_\theta \right)
    - \frac{|\bs{v}|^2}{2\,\sigma^2} \right] \\
\end{equation}
The integral over radial velocity is simple and can be performed without
considering any other terms because uncertainties in these measurements will
not, in general, be covariant with the Gaia data. The integrand over tangential
velocities can then be written
\begin{equation}
  L(v_\alpha, v_\delta, d) \, p(v_\alpha, v_\delta) = \tilde{\mathcal{Z}} \,
    \exp \left[ -\frac{1}{2}\,\transp{\left(\bs{x}' - \bs{x}'_\theta \right)} \,
    \matrx{V}' \,
    \left(\bs{x}' - \bs{x}'_\theta \right)
    - \frac{(v_\alpha^2 + v_\delta^2)}{2\,\sigma^2} \right]
    \label{eq:vtan-integrand}
\end{equation}
where the primed variables indicate that they are in the sub-space of just the
astrometric quantities, the constant pre-factor now includes the integral over
radial velocity
\begin{align}
  \tilde{\mathcal{Z}} = \mathcal{Z} \,
    \exp\left[-\frac{1}{2}\frac{v_r^2}{\sigma_{v_r}^2 + \sigma^2}\right] \,
    \left[ 2\pi \, (\sigma_{v_r}^2 + \sigma^2) \right]^{1/2}
\end{align}
and $\sigma_{v_r}$ is the observational uncertainty of the radial velocity
measurement. To perform the integral over tangential velocities, we partition
the inverse variance matrix and vectors so that
\begin{align}
  \bs{x}_A &= (\varpi)\\
  \bs{x}_B &= (\mu_\alpha, \mu_\delta)\\
  \bs{x}_{\theta, A} &= (d^{-1}) \\
  \bs{x}_{\theta, B} &= (d^{-1}\,v_\alpha, d^{-1}\,v_\delta) \\
  \matrx{V} &=
    \left(\begin{array}{c| c c}
      V_{11} & V_{12} & V_{13}\\\hline
      V_{12} & V_{22} & V_{23}\\
      V_{13} & V_{23} & V_{33}
    \end{array}\right)
    = \left(\begin{array}{c c}
      \matrx{V}_{AA} & \matrx{V}_{AB}\\
      \matrx{V}_{AB} & \matrx{V}_{BB}
    \end{array}\right)
\end{align}
\todo{Holy fuck that matrix is UGLY - fix with new version of AASTex 6}
and the argument of the Gaussian in \eqname~\ref{eq:vtan-integrand} can be
written
\begin{multline}
  -\frac{1}{2}\,[
  \transp{(\bs{x}_A - \bs{x}_{\theta, A})}\,
    \matrx{V}_{AA}\,(\bs{x}_A - \bs{x}_{\theta, A})\\
  + \transp{(\bs{x}_B - \bs{x}_{\theta, B})}\,
    \matrx{V}_{BB}\,(\bs{x}_B - \bs{x}_{\theta, B})\\
  + 2\,\transp{(\bs{x}_A - \bs{x}_{\theta, A})}\,
    \matrx{V}_{AB}\,(\bs{x}_B - \bs{x}_{\theta, B})]
\end{multline}
\todo{prior in above}

>>>>

% define macros for text
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\figname}{Figure}
\newcommand{\eqname}{Equation}
\newcommand{\dr}{\acronym{DR1}}

% define macros for math
\newcommand{\given}{\,|\,}
\newcommand{\normal}{{\mathcal{N}}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transp}[1]{{#1}^{\!\mathsf{T}}}
\newcommand{\inv}[1]{{#1}^{-1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\vperp}{\bs{v}^\perp}
\newcommand{\propm}{\bs{\mu}}
\newcommand{\matrx}[1]{\mathbf{#1}}
\newcommand{\kms}{\rm km~s^{-1}}
\newcommand{\data}{\mathrm{data}}
\newcommand{\snr}{[S/N]_\varpi}

@@

\title{Co-moving stars in \textsl{Gaia DR1}}
\author{Adrian M. Price-Whelan\altaffilmark{\pu,\adrn},
        Semyeong Oh\altaffilmark{\pu},
        David W. Hogg\altaffilmark{\ccpp,\mpia},
        Timothy D. Morton\altaffilmark{\pu},
        David N. Spergel\altaffilmark{\pu,\cca}
}

@@

\todo{Some bullshit about binary stars...}

@@

The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
Solution (TGAS), released as a part of Data Release 1 of the Gaia mission
(\citealt{many}). TGAS contains astrometric measurements (sky position,
parallax, and proper motions) and associated covariance matrices for a large
fraction of the \project{Tycho-2} catalog (\citealt{tycho2}) with median
astrometric precision equivalent to the \project{Hipparcos} catalog ($\approx
0.3~{\rm mas}$; \citealt{hipparcos}). In terms of parallax signal-to-noise
($\snr = \varpi/\sigma_\varpi$), the TGAS catalog contains 1240114 stars
with $\snr > 4$ and 619618 stars with $\snr > 8$. The increase in precision over
Tycho-2 and increase in volume over Hipparcos makes the TGAS catalog a powerful
data set to search for widely-separated, co-moving stars.

In the sections that follow, we construct a statistical model that use the TGAS
data and associated covariances to handle the non-trivial uncertainties [...].
Here, we instead visualize simple quantities constructed directly from the
astrometric measurements to emphasize the quality of the data. For each star
with $\snr > 8$, we use the reported parallax and parallax error to compute a
distance corrected for the Lutz-Kelker bias (\citealt{lutzkelker}):
\begin{equation}
  d = 1000 \, \left[\frac{\varpi}{2} \, \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
\end{equation}

@@
